Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-18 nips_2018.tex
greek lettering for proposed expressions
bunches of changes

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usepackage{circuitikz}
\usetikzlibrary{fit,bayesnet}
\usepackage{dsfont}
\usepackage{amsmath}

@@

We apply \system to three domains:
list processing; text editing (in the style of FlashFill~\cite{gulwani2011automating}); and symbolic regression.
 For each of these we initially provide a generic
 set of programming primitives.
Our algorithm then constructs
its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{table}[hb]
\makebox[\textwidth][c]{
\scriptsize
\renewcommand\code\texttt
\renewcommand\codechar[1]{\texttt{"#1"}}
\begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}}\toprule
  {\normalsize Lists}&{\normalsize Text}&{\normalsize Symbolic Regression}\\\midrule
  \begin{tabular}{c}
      % \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
      % \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
      % \code{$f(x) = $}\code{($f_0$ nil $x$)}\\\\
      % \code{[99\, 4\, 2]}$\to $\code{99}\\
      % \code{[2\, 1\, 4\, 6]}$\to $\code{6}\\
      % \code{[8\, 7\, 7\, 7]}$\to $\code{8}\\
      % \code{$f(\ell) = $}\code{($f_0$ $\ell$)}\\\\
      \code{[1]}$\to $\code{[1\, 3]}\\
      \code{[7\, 14\, 9]}$\to $\code{[7\, 14\, 9\, 3]}\\
      \code{$f(\ell) = $}\code{($f_1$ $\ell$ 3)}\\\\
      \code{[7\, 3\, 14\, 6\, 9]}$\to $\code{false}\\
      \code{[9\, 4\, 3\, 4]}$\to $\code{true}\\
      \code{$f(\ell) = $}\code{($f_2$ $\ell$ 4)}
  \end{tabular}&
  \begin{tabular}{c}
    +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
    +83 973-831$\to $83.973.831\\
    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} }\\
    \hspace{1cm}\code{($f_0$ \codechar{.} \codechar{ } (cdr s)))}\\\\
Temple Anna H $\to $TAH\\
Lara Gregori$\to $LG\\
$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}\\

@@

  \begin{tabular}{l}
    % $f_0($\code{r}$,\ell) \,=\, $\code{(foldr $\ell$ r ($\lambda$ (x a)}\\
    % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(foldr}}\code{(cons (index (length a) $\ell$) a)))}\\
      % \hspace{0.5cm}($f_1$: \emph{Get the largest number})\\
    % $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
      % \hspace{0.5cm}($f_1$: \emph{Get the largest number in $\ell$})\\
    %NOTE: this is the actual invention, but I removed a redundant lambda
    % below: $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ ($\lambda$ (x a) (cons x a)))}\\
    $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}\\
    \vspace{2ex}
      \hspace{0.5cm}($f_0$: \emph{Concatenate }\code{r}\emph{ with $\ell$})\\
    $f_1(\ell,$\code{k}$) \,=\, $\code{($f_0$ (cons k nil) $\ell$)}\\
    \vspace{2ex}
      \hspace{0.5cm}($f_1$: \emph{Append }\code{k}\emph{ to $\ell$})\\
    $f_2(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}\\
    \phantom{$f_1(\ell,$\code{k}}
    \code{($\lambda$ (x a) (if a a (= k x))))}\\
      \hspace{0.5cm}($f_2$: \emph{Whether $\ell$ contains }\code{k})\\
  \end{tabular}&

  \begin{tabular}{l}
          $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x)}\\\hspace{1cm}\code{ (if (= x a) b x)) s)}\\
      \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a)}\\\hspace{1.1cm}\code{ (cdr (if (= c x) s a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car }\\\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}\\
        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})\\
    \end{tabular}&

  \begin{tabular}{l}
    $f_0($\code{x}$)\,=\,$\code{(+ x real)}\\
    $f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} \\
    $f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}\\
    $f_3($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_2$\code{ x)))}\\
    $f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}\\
    \hspace{0.5cm}\emph{($f_4$: 4th order polynomial)}\\
        $f_5($\code{x}$)\,=\,$\code{(/ real x)}\\
    $f_6($\code{x}$)\,=\,$\code{($f_4$ ($f_0$ x))} \\
    \hspace{0.5cm}\emph{($f_6$: rational function)}\\

  \end{tabular}
  \\\bottomrule\\
\end{tabular}}
\caption{Top: Example tasks from each domain, each followed by the programs \system discovers for them. Bottom: Subset of learned DSL. Notice that learned DSL primitives can call each other.}\label{initialExampleDSL}
\end{table}

@@

Formally, $q$ should approximate the true posteriors over programs: minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
 equivalently maximizing $  \expect[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) ]$,
%% \begin{equation*}
%%   \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
%% \end{equation*}
 where the expectation is taken over tasks. Taking this expectation over the empirical distribution of tasks gives self-supervised training; taking it over samples from the generative model gives  Helmholtz-machine style training.
 The  objective for a recognition model ($\mathcal{L}_{\text{RM}}$) combines the Helmholtz machine ($\mathcal{L}_{\text{HM}}$) and self supervised ($\mathcal{L}_{\text{SS}}$) objectives, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{SS} + \mathcal{L}_\text{HM}$:
\begin{align}\nonumber
\mathcal{L}_{\text{HM}} = \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\quad
\mathcal{L}_{\text{SS}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
\end{align}
\begin{comment}
  The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network $q$.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $q$ on those samples.
  \end{comment}
Evaluating $\mathcal{L}_{\text{HM}}$ involves sampling programs from
the current DSL, running them to get their outputs,
and then training $q$ to regress from the input/outputs to the program.
Since these programs map inputs to outputs,
we need to sample the inputs as well.
Our solution is to sample the inputs
from the empirical observed distribution of inputs in $X$.

@@

\begin{wrapfigure}{R}{0.6\textwidth}
  \begin{minipage}{0.6\textwidth}    
    \begin{algorithm}[H]
      \caption{The \system Algorithm}
      \label{mainAlgorithm}
      \begin{algorithmic}
        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
        \FOR{$i=1$ {\bfseries to} $I$}
        %     \FOR{$x:\tau\in X$}
        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Search})}
        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Search)}}

@@

\begin{table}[b]
\centering  \begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (a b) (foldr b (cons "," a)}\\
      \phantom{\code{($\lambda$ }}\code{($\lambda$ (x z) (cons x z))))}\\
      \code{($\lambda$ (a b) (foldr a b }\\
      \phantom{\code{($\lambda$ }}\code{($\lambda$ (x z) (cons x z))))}
    \end{tabular}&
    \begin{tabular}{l}
      \code{(foldr a b ($\lambda$ (x z) }\\
      \phantom{\code{(foldr a b }}\code{(cons x z)))}
    \end{tabular}\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
      \end{tabular}&
    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x a) b x))) s)}
      \end{tabular}
\\\bottomrule\\
  \end{tabular}
  \caption{The DSL induction algorithm proposes fragments of programs to add to the DSL.
    These fragments are taken from programs in the frontiers (left column). Here, the proposed subexpression (top) appends two lists or (bottom) performs character substitutions.} 
  \label{fragmentExample}
\end{table}

@@

We automatically generated 109 text editing tasks with 4 input/output examples each
%(Fig.~\ref{exampleTextProblem}).
At first, \system cannot find any correct programs for most of the tasks.
After three iterations, it assembles a DSL (Fig.~\ref{firstPageFigure} \& center of Tbl.~\ref{initialExampleDSL}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks. The learned DSL contains \textbf{X} new functions, listed in the supplement.

@@

We compare with ablations of our model on held out tasks.
Each of these ablations is chosen to approximate
a prior approach in the program learning literature:
\\\noindent \textbf{Ours (no NN)}, which lesions the recognition model.
\\\noindent \textbf{RF/DC}, which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
a recognition model only from samples from the fixed generative model.
This is equivalent to our algorithm with $\lambda = \infty$ (Sec.~\ref{grammarInductionSection})
and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (Sec.~\ref{recognitionSection}).
We call this baseline RF/DC because
this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
We can not compare directly with these systems,
because they are engineered for one specific domain, and 
do not have publicly available code and datasets.
\\\noindent \textbf{EC}, which lesions the recognition model and modifies  the DSL learner to model the library learning technique of  EC~\cite{Dechter:2013:BLV:2540128.2540316}.
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first search step.

For each domain,
we are interested both in how many tasks the
agent can solve and how quickly it can find those solutions.
%% For symbolic regression, we also care about the quality
%% of the solution, as measured by the likelihood model $\probability[x|p]$,
%% e.g. did the agent correctly explain a linear function
%% using two numbers, or did it introduce extraneous parameters?
Tbl.~\ref{baselineComparisons}
compares our model against these baselines.
Our full model consistently
improves on the baselines,
sometimes dramatically (text editing and symbolic regression).
The recognition model 
consistently increases the number of solved held-out tasks,
and lesioning it also slows down the convergence of the algorithm,

taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
This supports a view of the recognition model as a way of amortizing the cost of searching for programs.

\begin{wraptable}{r}{0.6\textwidth}
%% This radically shrinks the table
\tabcolsep=4pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{lllllll}
  \toprule& Ours& \multicolumn{1}{c}{\begin{tabular}{c}
    Ours\\(no NN)
    \end{tabular}}& EC&RF/DC & PCFG & Enum

@@

  \\\bottomrule
  \end{tabular}
\caption{\% solved before timeout. Solve time: averaged over solved
  tasks.  RF/DC: trained like RobustFill/DeepCoder.  PCFG: model w/o
  structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}

>>>>

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usepackage{circuitikz}
\usetikzlibrary{fit,bayesnet}
\usepackage{dsfont}
\usepackage{amsmath}
\usetikzlibrary{trees}

@@

We apply \system to three domains:
list processing; text editing (in the style of FlashFill~\cite{gulwani2011automating}); and symbolic regression.
 For each of these we initially provide a generic
 set of programming primitives.
Our algorithm then constructs
its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{table}[hb]
\makebox[\textwidth][c]{
  \scriptsize
  \tabcolsep=4pt
\renewcommand\code\texttt
\renewcommand\codechar[1]{\texttt{"#1"}}
\newcommand{\helpSize}{0.25cm}
\begin{tabular}{>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}}\toprule
  {\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
  {  \tabcolsep=4pt \begin{tabular}{cc}
  \begin{tabular}{c}
      \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
      \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
      \code{$f(x) = $}\code{($f_0$ nil $x$)}\\\\
      \code{[2\, 7\, 8\, 1]}$\to $\code{8}\\
      \code{[3\, 19\, 14]}$\to $\code{19}\\
      \code{$f(\ell) = $}\code{($f_0$ $\ell$)}
      %% \code{[1]}$\to $\code{[1\, 3]}\\
      %% \code{[7\, 14\, 9]}$\to $\code{[7\, 14\, 9\, 3]}\\
      %% \code{$f(\ell) = $}\code{($f_1$ $\ell$ 3)}\\\\
  \end{tabular}&
  \begin{tabular}{c}
          \code{[7\, 3]}$\to $\code{\#f}\\
          \code{[9\, 4\, 4]}$\to $\code{\#t}\\
          \code{[]}$\to $\code{\#f}\\
          \code{[4\, 7\, 3]}$\to $\code{\#t}\\
      \code{$f(\ell) = $}\code{($f_2$ $\ell$ 4)}
    \end{tabular}
  \end{tabular}}
  &
  \begin{tabular}{c}
    +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
    +83 973-831$\to $83.973.831\\
    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} }\\
    \hspace{1cm}\code{($f_0$ \codechar{.} \codechar{ } (cdr s)))}\\\\
Temple Anna H $\to $TAH\\
Lara Gregori$\to $LG\\
$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}\\

@@

  \begin{tabular}{l}
    % $f_0($\code{r}$,\ell) \,=\, $\code{(foldr $\ell$ r ($\lambda$ (x a)}\\
    % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(foldr}}\code{(cons (index (length a) $\ell$) a)))}\\
      % \hspace{\helpSize}($f_1$: \emph{Get the largest number})\\
    % $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
      % \hspace{\helpSize}($f_1$: \emph{Get the largest number in $\ell$})\\
    %NOTE: this is the actual invention, but I removed a redundant lambda
    % below: $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ ($\lambda$ (x a) (cons x a)))}\\
    $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}\\
    \vspace{2ex}
      \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
    $f_1(\ell,$\code{k}$) \,=\, $\code{($f_0$ (cons k nil) $\ell$)}\\
    \vspace{2ex}
      \hspace{\helpSize}($f_1$: \emph{Append }\code{k}\emph{ to $\ell$})\\
    $f_2(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}\\
    \phantom{$f_1(\ell,$\code{k}}
    \code{($\lambda$ (x a) (if a a (= k x))))}\\
      \hspace{\helpSize}($f_2$: \emph{Whether $\ell$ contains }\code{k})\\
  \end{tabular}&

  \begin{tabular}{l}
          $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x)}\\\hspace{1cm}\code{ (if (= x a) b x)) s)}\\
      \hspace{\helpSize}($f_0$: \emph{Performs character substitution)}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a)}\\\hspace{1.1cm}\code{ (cdr (if (= c x) s a))))}\\
        \hspace{\helpSize}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car }\\\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}\\
        \hspace{\helpSize}($f_2$: \emph{Abbreviates a sequence of words})\\
    \end{tabular}&

  \begin{tabular}{l}
    $f_0($\code{x}$)\,=\,$\code{(+ x real)}\\
    $f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} \\
    $f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}\\
    $f_3($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_2$\code{ x)))}\\
    $f_4($\code{x}$)\,=\,$\code{($f_0$ (* x (}$f_3$\code{ x)))}\\
    \hspace{\helpSize}\emph{($f_4$: 4th order polynomial)}\\
        $f_5($\code{x}$)\,=\,$\code{(/ real x)}\\
    $f_6($\code{x}$)\,=\,$\code{($f_4$ ($f_0$ x))} \\
    \hspace{\helpSize}\emph{($f_6$: rational function)}\\

  \end{tabular}
  \\\bottomrule\\
\end{tabular}}
\caption{Top: Tasks from each domain, each followed by the programs \system discovers for them. Bottom: Several examples from learned DSL. Notice that learned DSL primitives can call each other.}\label{initialExampleDSL}
\end{table}

@@

Formally, $q$ should approximate the true posteriors over programs: minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|q(p|x) \right) \right]$,
 equivalently maximizing $  \expect[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) ]$,
%% \begin{equation*}
%%   \expect\left[\sum_p\probability[p|x,\mathcal{D},\theta]\log q(p|x) \right]
%% \end{equation*}
 where the expectation is taken over tasks. Taking this expectation over the empirical distribution of tasks gives self-supervised training; taking it over samples from the generative model gives  Helmholtz-machine style training.
 The  objective for a recognition model ($\mathcal{L}_{\text{RM}}$) combines the Helmholtz machine ($\mathcal{L}_{\text{HM}}$) and self supervised ($\mathcal{L}_{\text{SS}}$) objectives, $\mathcal{L}_{\text{RM}} = \mathcal{L}_\text{SS} + \mathcal{L}_\text{HM}$:
\begin{align}\nonumber
\mathcal{L}_{\text{HM}} = \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log q(p|x)\right]\quad
\mathcal{L}_{\text{SS}} = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
  \frac{\probability\left[x,p|\mathcal{D},\theta \right]}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\log q(p|x)\right]
\end{align}
\begin{comment}
  The $\mathcal{L}_{\text{HM}}$ objective is essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network $q$.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $q$ on those samples.

Evaluating $\mathcal{L}_{\text{HM}}$ involves sampling programs from
the current DSL, running them to get their outputs,
and then training $q$ to regress from the input/outputs to the program.
Since these programs map inputs to outputs,
we need to sample the inputs as well.
Our solution is to sample the inputs
from the empirical observed distribution of inputs in $X$.
  \end{comment}

@@

\begin{wrapfigure}{R}{0.6\textwidth}
  \begin{minipage}{0.6\textwidth}    
    \begin{algorithm}[H]
      \caption{The \system Algorithm}
      \label{mainAlgorithm}
      \begin{algorithmic}
        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
        \FOR{$i=1$ {\bfseries to} $I$}
        %     \FOR{$x:\tau\in X$}
        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in
        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
        \footnotesize{\hspace{0.1cm}(\textbf{Search})}
        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in
        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
        \footnotesize{\hspace{0.15cm}\textbf{(Search)}}

@@

\begin{figure}[b!]\tabcolsep=2pt
\centering  \begin{minipage}[c]{0.2\textwidth}
  \begin{tikzpicture}[scale=0.7]
    %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
    %%     \footnotesize Program:\\
    %%     \code{($\lambda$ (x) (+ (- x) 1))}
    %% \end{tabular}};
    %\node at (,0.5) {\code{cons}};
    \node [rotate=90] at (-2.3,-0.5) {\small program};

          \node(l1) at (0,0) {};
  \node[color=red](p1) at (-1,-1) {\code{+}};
  \node[color=red](n1) at (0.7,-0.9) {\code{1}};
  \node(x1) at (0,-1) {\code{1}};
  \draw[color=red] (l1.south) -- (p1.north);
  \draw[color=red] (l1.south) -- (n1.north);
  \draw[color=red] (-0.5,-0.5) -- (x1.north);

  \node(t) at (-0.5,0.5) {};
  \draw (l1.south) -- (t.south);
  \node(c) at (-1.5,-0.2) {\code{cons}};
  \draw (t.south) -- (c.north);

%    \draw  (l1.south) -- (-0.5,0.5);

  %% \node(c) at (-0.5,-1.5) {\code{-}};
  %% \node(z) at (0.5,-1.5) {\code{x}};

  %% \draw (0,-1) -- (c.north);
  %% \draw (0,-1) -- (z.north);

  \begin{scope}[shift={(0,-2.5)}]
      \node(l1) at (0,0) {};
  \node[color=red](p1) at (-1,-1) {\code{+}};
  \node[color=red](n1) at (0.7,-0.9) {\code{1}};
  %\node(x1) at (0,-1) {};
  \draw[color=red] (l1.south) -- (p1.north);
  \draw[color=red] (l1.south) -- (n1.north);
  \draw[color=red] (-0.5,-0.5) -- (0,-1);

  \node(c) at (-0.5,-1.5) {\code{car}};
  \node(z) at (0.5,-1.5) {\code{z}};

  \draw (0,-1) -- (c.north);
  \draw (0,-1) -- (z.north);

  \node [rotate=90] at (-2.3,-0.7) {\small program};

  \end{scope}

\begin{scope}[shift={(0,-5)}]
  \node[red](p1) at (-1,-1) {\code{+}};
  \node[red](n1) at (0.8,-0.7) {\code{1}};
  \node[red](a) at (0,-1) {\code{ }};
  %\node(x1) at (0,-1) {};
  \draw[red] (0,0) -- (p1.north);
  \draw[red] (0,0) -- (n1.north);
  \draw[red] (-0.55,-0.4) -- (a.north);
  \node [rotate=90] at (-2.3,-0.7) {\small fragment};

  \end{scope}

\end{tikzpicture}
  \end{minipage}
\hspace{0.1cm}\begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed $\lambda$-expression\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (- (length $\ell$) 1))))}\\
      \code{($\lambda$ ($\ell$) (map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{($\lambda$ ($\ell$) (map }}\code{(range (+ 1 1))))}\\
    \end{tabular}&
    \begin{tabular}{l}
      \code{(map ($\lambda$ (x) (index x $\ell$))}\\
      \phantom{\code{(map)}}\code{(range $\alpha$))}\\
    \end{tabular}\\\midrule
    \begin{tabular}{l}
      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\
      \hspace{0.4cm}\code{ (if (= x '.') '-' x))) s)}\\
      \code{($\lambda$ (s) (map ($\lambda$ (x) }\\
      \hspace{0.4cm}\code{ (if (= x '-') ',' x))) s)}\\
      \end{tabular}&
    \begin{tabular}{l}      \code{($\lambda$ (s) (map ($\lambda$ (x)}\\\hspace{0.2cm}\code{   (if (= x $\alpha$) $\beta$ x))) s)}
      \end{tabular}
\\\bottomrule\\
\end{tabular}
\caption{\textbf{Left:} syntax trees of two programs sharing common structure, highlighted in {\color{red} red},
  from which we extract a fragment and add it to the DSL (bottom). \textbf{Right:} actual programs,
  from which we extract fragments that  (top) slice from the beginning of a list or (bottom) perform character substitutions.}

  \label{fragmentExample}
\end{figure}

@@

We automatically generated 109 text editing tasks with 4 input/output examples each
%(Fig.~\ref{exampleTextProblem}).
At first, \system cannot find any correct programs for most of the tasks.
After three iterations, it assembles a DSL (Fig.~\ref{firstPageFigure} \& center of Tbl.~\ref{initialExampleDSL}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks. The learned DSL contains 10 new functions, listed in the supplement.

@@

We compare with ablations of our model on held out tasks.
Each of these ablations is chosen to approximate
a prior approach in the program learning literature:
\\\noindent \textbf{No NN}, which lesions the recognition model.
\\\noindent \textbf{NPS}, which does not learn the DSL,
instead learning the recognition model
from samples drawn from the fixed DSL.
We call this NPS (Neural Program Synthesis)
because this setup is closest to how
RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
%% We can not compare directly with these systems,
%% because they are engineered for one specific domain, and 
%% do not have publicly available code and datasets.
\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
only add subexpressions of programs in the frontiers to the DSL. This is how prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first search step.

For each domain,
we are interested both in how many tasks the
agent can solve and how quickly it can find those solutions.
%% For symbolic regression, we also care about the quality
%% of the solution, as measured by the likelihood model $\probability[x|p]$,
%% e.g. did the agent correctly explain a linear function
%% using two numbers, or did it introduce extraneous parameters?
Tbl.~\ref{baselineComparisons}
compares our model against these alternatives.
Our full model consistently
improves on the baselines,
sometimes dramatically (text editing and symbolic regression).
The recognition model 
consistently increases the number of solved held-out tasks,
and lesioning it also slows down the convergence of the algorithm,
taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
This supports a view of the recognition model as a way of amortizing the cost of searching for programs.

\begin{wraptable}{r}{0.6\textwidth}
%% This radically shrinks the table
\tabcolsep=4pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{lllllll}
  \toprule& Ours& 
    No NN
& SE&NPS & PCFG & Enum

@@

  \\\bottomrule
  \end{tabular}
\caption{\% solved before timeout. Solve time: averaged over solved
  tasks.  SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
  structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
